---
title: "ACTL1101 Assignment Part A"
author: "Meharpreet Kaur (z5418597)"
date: "2024 T2"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Algorithmic Trading Strategy

## Introduction

In this assignment, you will develop an algorithmic trading strategy by incorporating financial metrics to evaluate its profitability. This exercise simulates a real-world scenario where you, as part of a financial technology team, need to present an improved version of a trading algorithm that not only executes trades but also calculates and reports on the financial performance of those trades.

## Background

Following a successful presentation to the Board of Directors, you have been tasked by the Trading Strategies Team to modify your trading algorithm. This modification should include tracking the costs and proceeds of trades to facilitate a deeper evaluation of the algorithm’s profitability, including calculating the Return on Investment (ROI).

After meeting with the Trading Strategies Team, you were asked to include costs, proceeds, and return on investments metrics to assess the profitability of your trading algorithm.

## Objectives

1. **Load and Prepare Data:** Open and run the starter code to create a DataFrame with stock closing data.

2. **Implement Trading Algorithm:** Create a simple trading algorithm based on daily price changes.

3. **Customize Trading Period:** Choose your entry and exit dates.

4. **Report Financial Performance:** Analyze and report the total profit or loss (P/L) and the ROI of the trading strategy.

5. **Implement a Trading Strategy:** Implement a trading strategy and analyze the total updated P/L and ROI. 

6. **Discussion:** Summarise your finding.


## Instructions

### Step 1: Data Loading

Start by running the provided code cells in the "Data Loading" section to generate a DataFrame containing AMD stock closing data. This will serve as the basis for your trading decisions. First, create a data frame named `amd_df` with the given closing prices and corresponding dates. 

```{r load-data}

# Load data from CSV file
amd_df <- read.csv("AMD.csv")

# Convert the date column to Date type and Adjusted Close as numeric
amd_df$date <- as.Date(amd_df$Date)
amd_df$close <- as.numeric(amd_df$Adj.Close)

amd_df <- amd_df[, c("date", "close")]
```


##Plotting the Data
Plot the closing prices over time to visualize the price movement.
```{r plot}
plot(amd_df$date, amd_df$close,'l')
```


## Step 2: Trading Algorithm
Implement the trading algorithm as per the instructions. You should initialize necessary variables, and loop through the dataframe to execute trades based on the set conditions.

- Initialize Columns: Start by ensuring dataframe has columns 'trade_type', 'costs_proceeds' and 'accumulated_shares'.
- Change the algorithm by modifying the loop to include the cost and proceeds metrics for buys of 100 shares. Make sure that the algorithm checks the following conditions and executes the strategy for each one:
  - If the previous price = 0, set 'trade_type' to 'buy', and set the 'costs_proceeds' column to the current share price multiplied by a `share_size` value of 100. Make sure to take the negative value of the expression so that the cost reflects money leaving an account. Finally, make sure to add the bought shares to an `accumulated_shares` variable.
  - Otherwise, if the price of the current day is less than that of the previous day, set the 'trade_type' to 'buy'. Set the 'costs_proceeds' to the current share price multiplied by a `share_size` value of 100.
  - You will not modify the algorithm for instances where the current day’s price is greater than the previous day’s price or when it is equal to the previous day’s price.
  - If this is the last day of trading, set the 'trade_type' to 'sell'. In this case, also set the 'costs_proceeds' column to the total number in the `accumulated_shares` variable multiplied by the price of the last day.



```{r trading}
# Initialize columns for trade type, cost/proceeds, and accumulated shares in amd_df
amd_df$trade_type <- NA # Initialize trade_type column
amd_df$trade_type[1] <- "buy" # Fix first variable
for (i in 2:nrow(amd_df)) {if(amd_df$close[i]<amd_df$close[i-1]){amd_df$trade_type[i] <- "buy"}else{amd_df$trade_type[i] <- ""}} #for the rows between 2 and nrow(amd_df) if the value of close was less than the value of the previous close than the cell will display a "buy". If the value was greater than or equal the cell will be empty
amd_df$trade_type[nrow(amd_df)] <- "sell" # Fix last variable
amd_df$costs_proceeds <- NA # Initialize costs_proceeds column
for (i in 1:nrow(amd_df)) {if(amd_df$trade_type[i] == "buy"){amd_df$costs_proceeds[i] <- -100*amd_df$close[i]}else{amd_df$costs_proceeds[i]<-0}} #for rows between 1 and nrow(amd_df) if the trade_type column displayed "buy" cost proceeds would be calculated at the 100 shares multiplied by the close price for the date. If trade_type was empty, cost proceeds were 0. 
#the final sell price requires the accumalted shares column will insert coding after 
amd_df$accumulated_shares <- 0 #Initialize accumalted_shares column
amd_df$accumulated_shares[1] <- 100 #fix initial purchase of 100 shares
for (i in 2:nrow(amd_df)) {if(amd_df$trade_type[i] == "buy"){amd_df$accumulated_shares[i] <- amd_df$accumulated_shares[i-1] + 100}else{amd_df$accumulated_shares[i] <-amd_df$accumulated_shares[i-1]}} #for rows between 2 and nrow(amd_df), if the trade_type column displayed a buy 100 shares were added to the value of the previous date. If the trade_type cell was empty, the accumalted shares only presneted the value from the previous date.
amd_df$costs_proceeds[nrow(amd_df)]<-amd_df$accumulated_shares[nrow(amd_df)]*amd_df$close[nrow(amd_df)] # to calculate final costs proceeded the close cost of the final day was multiplied by the total accumulated shares.
```


## Step 3: Customize Trading Period
- Define a trading period you wanted in the past five years 
```{r period}
# Fill your code here
start_date <- as.Date("2020-01-02")
end_date <- as.Date("2022-01-03")
new_amd <- amd_df[amd_df$date>=start_date & amd_df$date <=end_date, ]
new_amd$trade_type <- NA
new_amd$costs_proceeds <- NA
new_amd$accumulated_shares <- 0
new_amd$trade_type[1] <- "buy"
new_amd$trade_type[nrow(new_amd)] <- "sell"
for (i in 2:nrow(new_amd)) {if(new_amd$close[i]<new_amd$close[i-1]){new_amd$trade_type[i] <- "buy"}else{new_amd$trade_type[i] <- ""}}
new_amd$costs_proceeds <- NA
for (i in 1:nrow(new_amd)) {if(new_amd$trade_type[i] == "buy"){new_amd$costs_proceeds[i] <- -100*new_amd$close[i]}else{new_amd$costs_proceeds[i]<-0}}
new_amd$accumulated_shares <- 0 
new_amd$accumulated_shares[1] <- 100
for (i in 2:nrow(new_amd)) {if(new_amd$trade_type[i] == "buy"){new_amd$accumulated_shares[i] <- new_amd$accumulated_shares[i-1] + 100}else{new_amd$accumulated_shares[i] <-new_amd$accumulated_shares[i-1]}}
new_amd$costs_proceeds[nrow(new_amd)]<-new_amd$accumulated_shares[nrow(new_amd)]*new_amd$close[nrow(new_amd)]
#replicating step 2 for new_amd

```


## Step 4: Run Your Algorithm and Analyze Results
After running your algorithm, check if the trades were executed as expected. Calculate the total profit or loss and ROI from the trades.

- Total Profit/Loss Calculation: Calculate the total profit or loss from your trades. This should be the sum of all entries in the 'costs_proceeds' column of your dataframe. This column records the financial impact of each trade, reflecting money spent on buys as negative values and money gained from sells as positive values.
- Invested Capital: Calculate the total capital invested. This is equal to the sum of the 'costs_proceeds' values for all 'buy' transactions. Since these entries are negative (representing money spent), you should take the negative sum of these values to reflect the total amount invested.
- ROI Formula: $$\text{ROI} = \left( \frac{\text{Total Profit or Loss}}{\text{Total Capital Invested}} \right) \times 100$$

```{r}
# Fill your code here
total_profit <- sum(new_amd$costs_proceeds) #Total profit/loss calculation
capital_invested <- -sum(new_amd$costs_proceeds[new_amd$costs_proceeds<0]) #Total capital invested
roi <- total_profit/capital_invested #ROI
```

## Step 5: Profit-Taking Strategy or Stop-Loss Mechanisum (Choose 1)
- Option 1: Implement a profit-taking strategy that you sell half of your holdings if the price has increased by a certain percentage (e.g., 20%) from the average purchase price.
- Option 2: Implement a stop-loss mechanism in the trading strategy that you sell half of your holdings if the stock falls by a certain percentage (e.g., 20%) from the average purchase price. You don't need to buy 100 stocks on the days that the stop-loss mechanism is triggered.


```{r option 2}
# Fill your code here
new_amd$buy_sell <- NA #New variable representing trade_type
new_amd$acc_shares <- NA #New variable representing accumulated_shares
new_amd$capital <- NA #New variable representing costs_proceeded
average <- mean(new_amd$close[new_amd$trade_type == "buy"], na.rm=TRUE)
for (i in 1:nrow(new_amd)) {if(new_amd$close[i]>average*1.15){new_amd$buy_sell[i] <- "sell"}else{new_amd$buy_sell[i] <- "buy"}} #if the close price is greater than a 15% increase from the will sell. if not will buy.
new_amd$acc_shares[1] <- 100 #setting first variable -> buying 100 shares to begin
for (i in 2:nrow(new_amd)) {if(new_amd$buy_sell[i] == "buy"){new_amd$acc_shares[i] <- new_amd$acc_shares[i-1] + 100}else{new_amd$acc_shares[i] <-new_amd$acc_shares[i-1]/2}} #from last code if the close price is greater than 15% of the average price then will sell half its holdings. If buy_sell displays a "buy" then will add 100 more shares to the previous accumation. if it displays a sell then the previous accumlation will half.
for (i in 1:nrow(new_amd)) {if(new_amd$buy_sell[i] == "buy"){new_amd$capital[i] <- -100*new_amd$close[i]}else{new_amd$capital[i]<-new_amd$acc_shares[i]*new_amd$close[i]}} #if buy_sell displays a buy then the capital invested will be equal to -100 multiplied by the share price. if buy_sell displays a sell then the capital gained will be equal to half of the previous accumulation which would be the current value in acc_shares
```


## Step 6: Summarize Your Findings
- Did your P/L and ROI improve over your chosen period?
- Relate your results to a relevant market event and explain why these outcomes may have occurred.


```{r}
# Fill your code here and Disucss
TP <- sum(new_amd$capital)
CI <- -sum(new_amd$capital[new_amd$capital<0])
ROI <- TP/CI
TP_OG <- sum(amd_df$costs_proceeds)
CI_OG <- -sum(amd_df$costs_proceeds[amd_df$costs_proceeds<0])
ROI_OG <- TP_OG/CI_OG
print(ROI)
print(ROI_OG)
print(TP)
print(TP_OG)
plot(new_amd$date, new_amd$close,'l') #generating graph for trading period to better understand the data
#Discussion:  
#On Tuesday, December 8th, AMD launched “Radeon RX 6900 XT Graphics Card” which was designed to improve the quality of frames and resolution for high-intensity gaming. The following Tuesday the share price increased to over 15% of the average purchase. My strategy proposed to sell and occulated $1170296.00 as opposed to so profit for that day. In general, the original strategy proposed a higher ROI and profit.
```

Sample Discussion: On Wednesday, December 6, 2023, AMD CEO Lisa Su discussed a new graphics processor designed for AI servers, with Microsoft and Meta as committed users. The rise in AMD shares on the following Thursday suggests that investors believe in the chipmaker's upward potential and market expectations; My first strategy earned X dollars more than second strategy on this day, therefore providing a better ROI.
